---
title: "Customer_Segmentation"
author: "Tyler Mann"
date: "2023-05-23"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Customer Segmentation
## R Markdown

This report will document the process of conducting a customer segmentation analysis and follow the structure stated below:

### Types of Customer Segmentation Analyses:
1 - Demographic (age, gender, income)
2 - Behavioral (purchasing patterns, freq, AOV, churn)
3 - Psychographic - underlying motives, values, interests
4 - Lifestyle - lifestyles or life stages
5 - Channel Preference
6 - Purchase Journey - touchpoints or pipelines pts
7 - Customer Lifetime Value - high val vs low customers
8 - Product Usage - category grouping
9 - Customer Satisfaction - based on reviews
10 - Customer Preferences - based on prefs on pricing, promotions or delivery options. 

### Exploratory Data Analysis:
1 - Overview of Data
2 - Tx Volume
3 - Popular Products
4 - Customer Behavior
5 - Revenue Analysis
6 - Purchase Patterns
7 - Basket Size and Combos
8 - Customer Segmentation
9 - Data Quality and Completeness

### Checklist for Data Cleaning
1 - Validate Data Format
  Ensure correct data types and no bad formatting
2 - Handle Missing Values
  Decide on approach for missing values
    Mean, median, mode, regression, row deletion
3 - Remove Duplicates
4 - Cleanse Outliers
  Decide on approach to outlier values
    Remove or transform or otherwise
5 - Standardize Data
  Check if data needs standardization or normalization

### Checklist for EDA
1 - Identify Variables
2 - Univariate Analysis
  Analyze distribution, central tendency (mean, median), variability (stdev, range) and outliers / unusuals
3 - Bivariate Analysis
  Explore relationship between variables
  Scatterplots, correlation tables, cross-tabulations
4 - Missing Data Analysis
  Examine missingness and determine approach to resolution
5 - Outlier Detection
  Similar to above but for outliers
6 - Data Visualization
  Histograms, box plots, bar charts, scatter plots, etc.
7 - Feature Engineering
  New variables, scaling variables, deriving variables
8 - Multivariate Analysis
  PCA, factor analysis, cluster analysis, etc.
  1 - Dimension Reduction - PCA
  2 - Correlation and Covariance Analysis
  3 - Clustering Analysis
    K-means to identify natural groupings
  4 - Discriminant Analysis - LDA for classification
9 - Statistical Modeling
  Regression analysis, hyp testing, ANOVA
  

### If translated into a client deck:
1 - Introduction
  Introduce purpose and context
  Brief overview of business in question
  Explanation of importance of analysis
2 - Objective and Methodology
  State the specific objective of the analysis
  (What type of customer segmentation)
  Describe methodology
    clustering algorithms, decision trees, statistical techniques
  Briefly explain data sources, variables used and preprocessing
3 - Data Exploration
  Present key findings from EDA
  Highlight important insights about behavior, preferences or patterns
  Use visualizations or summary statistics to support findings
4 -  Segmentation Results
  Present results of segmentation analysis
  Describe the identified segments and their characteristics
  Display visual representations of segments such as cluster plots or decision trees to help visualize differences
5 - Segment Profiles
  Provide detailed profiles of each segment
  Discuss unique traits/behaviors/preferences of each segment
  Include relevant metrics or stats that distinguish segments
6 - Segment-specific insights and Recommendations
  Share actionable insights and recs for each segment
  Highlight opportunities for targeted marketing, personalized experiences or product/service customization
  Discuss strategies to improve customer satisfaction, retention or engagement for each segment
7 - Summary and Conclusion
  Recap the main findings of the customer segmentation analysis
  Emphasize the value of the insights gained and their potential impact on business outcomes
  Conclude with call to action or next steps for leveraging insights
8 - Q&A and Discussion
  Allocate time for questions, comments and discussion
  Be prepared to explain any technical aspects, assumptions or limitations of analysis
  
Statistical Tests
1 - T-test
    comparing means between segments
    example - time spent in shopping cart btwn groups
2 - ANOVA
    analysis of variance
    determine if signif differences in variability btwn groups
3 - chi square
    with categorical vars
    assess if signif association btwn cat vars across diff segs
4 - propensity score matching
    compare effect of specific treatment 
    match individuals in different groups with similar propensity scores for more reliable cross-seg comparison
  
### Data Import

```{r data}
library(data.table)
library(lubridate)
library(dplyr)
#install.packages("summarytools")
library(summarytools)
library(tidyverse)
library(ggplot2)

customer_data <- read_csv("Customer_Dataset_Data.csv")


```

## Data Review

Region is number range 1-5, probably fine

TownSize also a number range 1-5, has 2 NAs

Gender is a character... can fix

Age is a number, range 18 - 79, should be considered for segmentation

Education ranges from 6 to 23

JobCat is a character

Union Member is character, could be binary?

Retired could be converted to binary

HHIncome needs to be converted to number AND validated afterwards

Look into the ...15 column

LoanDefault could be binary

MaritalStatus could be binary

There are some more NAs after this Var!

HomeOwner could be binary and has NAs

CarValue could be number

CreditSpendMonth could be number

VoiceLastMonth == DataOverTenure could ALL be converted into numerical shape

Some of the following variables appear to be character cols that could be converted into Binary variables


```{r, echo=FALSE}

summary(customer_data)
lapply(customer_data, class)
head(customer_data)
#dim(customer_data)
##[1] 5000   60

```

## Data Cleaning

### Data Types

```{r}
colnames(customer_data)[0:14]
# [1] "CustomerID"        "Region"            "TownSize"          "Gender"            "Age"               "EducationYears"   
# [7] "JobCategory"       "UnionMember"       "EmploymentLength"  "Retired"           "HHIncome"          "DebtToIncomeRatio"
# [13] "CreditDebt"        "OtherDebt"
summary(customer_data)[,0:14]
summary(CD)[,0:14]
cdclean <- CD
nona <- CD[complete.cases(CD),]

cdclean$HHIncome <- gsub("$","",as.character(cdclean$HHIncome))
as.numeric(cdclean$HHIncome)
gsub(".00","",as.character(cdclean$HHIncome))

### EMOVE THE $ sign from HHINcome and then convert to numeric
newHH <- str_sub(cdclean$HHIncome,2,-4)
newHH <- gsub(",","",as.character(newHH))
newnumHH <- as.numeric(newHH)
newnumHH

nona <- transform(nona, HHIncome = as.numeric(HHIncome))
summary(nona)[,0:14]

colnames(customer_data)[15:29]
# [1] "...15"             "LoanDefault"       "MaritalStatus"     "HouseholdSize"     "NumberPets"        "NumberCats"       
# [7] "NumberDogs"        "NumberBirds"       "HomeOwner"         "CarsOwned"         "CarOwnership"      "CarBrand"         
# [13] "CarValue"          "CommuteTime"       "PoliticalPartyMem"
summary(customer_data)[,15:29]

colnames(customer_data)[30:44]
#[1] "Votes"               "CreditCard"          "CardTenure"          "CardItemsMonthly"    "CardSpendMonth"     
# [6] "ActiveLifestyle"     "PhoneCoTenure"       "VoiceLastMonth"      "VoiceOverTenure"     "EquipmentRental"    
# [11] "EquipmentLastMonth"  "EquipmentOverTenure" "CallingCard"         "WirelessData"        "DataLastMonth"   
summary(customer_data)[,30:44]

colnames(customer_data)[45:60]
# [1] "DataOverTenure"   "Multiline"        "VM"               "Pager"            "Internet"         "CallerID"        
# [7] "CallWait"         "CallForward"      "ThreeWayCalling"  "EBilling"         "TVWatchingHours"  "OwnsPC"          
# [13] "OwnsMobileDevice" "OwnsGameSystem"   "OwnsFax"          "NewsSubscriber"  
summary(customer_data)[,45:60]


```


### Missing Values

```{r}
# DROP ...15
CD <- customer_data[,c(1:14,16:60)]
## SKIP TO NONA
nona <- CD[complete.cases(CD),]


# Select all the rows with NAs as a subset of total dataset
CD_NA <- CD[!complete.cases(CD),]
colnames(CD_NA)

sapply(CD_NA, function(x) sum(is.na(x)))
is.na(CD_NA)
which(colSums(is.na(CD_NA))>0)

freq(CD_NA, var = Gender)

# DROP Number of Pets Variables
cd2 <- CD[,c(1:17,22:59)]
sapply(cd2, function(x) sum(is.na(x)))
which(colSums(is.na(cd2))>0)
freq(CD, var = TownSize) # MEAN REPLACE
freq(cd2, var = Gender) # DROP
freq(cd2, var = JobCategory) # MODE REPLACE
freq(cd2, var = HouseholdSize) # MEAN REPLACE
freq(cd2, var = HomeOwner) # MEAN REPLACE
freq(cd2, var = CommuteTime) # MEAN REPLACE



cd_new <- cd2

CD$TownSize[is.na(CD$TownSize)] <- mean(CD$TownSize, na.rm = TRUE)

cd_new$TownSize[is.na(cd2$TownSize)] <- mean(cd2$TownSize)
freq(cd_new, var = TownSize) # MEAN REPLACE
cd_new$TownSize[is.na(cd_new$JobCategory)] <- mode(cd_new$JobCategory)
cd_new$TownSize[is.na(cd_new$HouseholdSize)] <- mean(cd_new$HouseholdSize)
cd_new$TownSize[is.na(cd_new$HomeOwner)] <- mean(cd_new$HomeOwner)
cd_new$TownSize[is.na(cd_new$CommuteTime)] <- mean(cd_new$CommuteTime)

sapply(cd_new, function(x) sum(is.na(x)))
which(colSums(is.na(cd_new))>0)

cd3 <- cd2[!is.na(cd2$Gender),] ## 4967 obs
sapply(cd3, function(x) sum(is.na(x)))

cd4 <- cd3[!is.na(cd3$JobCategory),] # 4952 obs
sapply(cd4, function(x) sum(is.na(x)))

#####
nona <- CD[complete.cases(CD),]
#####


# https://www.statology.org/impute-missing-values-in-r/

df$col[is.na(df$col)] <- mean(df$col, na.rm=TRUE)

for(i in 1:ncol(df)) {
  df[ , i][is.na(df[ , i])] <- mean(df[ , i], na.rm=TRUE)
}

```


### Duplicate Values

```{r}



```

### Outliers

```{r}



```

### Standardize

```{r}



```




### SummaryTools()
```{r}
## SummaryTools()
#freq()
  #descr()
  #ctable()
  #compareGroups()
  #descrTable()
  #freq.poly()
  #dfSummary()

```

## Visualize

```{r}

#create histogram of values for price
ggplot(data=diamonds, aes(x=price)) +
  geom_histogram(fill="steelblue", color="black") +
  ggtitle("Histogram of Price Values")

#create scatterplot of carat vs. price, using cut as color variable
ggplot(data=diamonds, aes(x=carat, y=price, color=cut)) + 
  geom_point()

#create scatterplot of price, grouped by cut
ggplot(data=diamonds, aes(x=cut, y=price)) + 
  geom_boxplot(fill="steelblue")

```


# Multivariate Analysis

## Correlation Matrix

```{r}

#cor()
round(cor(diamonds[c('carat', 'depth', 'table', 'price', 'x', 'y', 'z')]), 2)

```


