---
title: "Customer_Segmentation"
author: "Tyler Mann"
date: "2023-05-23"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Customer Segmentation
## R Markdown

This report will document the process of conducting a customer segmentation analysis and follow the structure stated below:

If translated into a client deck:
1 - Introduction
  Introduce purpose and context
  Brief overview of business in question
  Explanation of importance of analysis
2 - Objective and Methodology
  State the specific objective of the analysis
  (What type of customer segmentation)
  Describe methodology
    clustering algorithms, decision trees, statistical techniques
  Briefly explain data sources, variables used and preprocessing
3 - Data Exploration
  Present key findings from EDA
  Highlight important insights about behavior, preferences or patterns
  Use visualizations or summary statistics to support findings
4 -  Segmentation Results
  Present results of segmentation analysis
  Describe the identified segments and their characteristics
  Display visual representations of segments such as cluster plots or decision trees to help visualize differences
5 - Segment Profiles
  Provide detailed profiles of each segment
  Discuss unique traits/behaviors/preferences of each segment
  Include relevant metrics or stats that distinguish segments
6 - Segment-specific insights and Recommendations
  Share actionable insights and recs for each segment
  Highlight opportunities for targeted marketing, personalized experiences or product/service customization
  Discuss strategies to improve customer satisfaction, retention or engagement for each segment
7 - Summary and Conclusion
  Recap the main findings of the customer segmentation analysis
  Emphasize the value of the insights gained and their potential impact on business outcomes
  Conclude with call to action or next steps for leveraging insights
8 - Q&A and Discussion
  Allocate time for questions, comments and discussion
  Be prepared to explain any technical aspects, assumptions or limitations of analysis
  


```{r data}



```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
#Plot of Pressure and second! test of Git
plot(pressure)
```


